Fast and optimal centroiding of faint stars

### 2015-03-09 ms/figure.tex
after word rearrangement and adding moments!

A common practice in astronomy is taking imaging data, and then finding the coordinates
of various light sources across the sky. Finding accuarate estimate of the center of point
sources, convolved with telescope point spread function (and atmospheric PSF in case of
ground based telescopes), and also pixel response function, is crucial to further steps of
astronomical image processing. For instance, proper measurement of the shapes of galaxies
requires interpolating the PSF from the position of sparsely located stars across the
image to the positions of galaxies. The accuracy of the PSF estimation therefore,
relies on how accurately we know the centroid of stars.

Ideally, we want a centroiding procedure that provides measurements as accurate as possible,
without putting a huge computational burden on the photometric pipeline.
Reducing the computational cost becomes even more important in large surveys,
where we want to estimate the centroid of thousands of point sources detected
on the telescope's focal plane.

Reaching the optimal measurement of the centroid of stars however, is limited
by the lack of knowledge about the exact shape of the PSF, and also presence of noise;
both sky noise and CCD readout noise. Thus, of particular interest is devising a fast method that returns
optimal estimates of centroids in a realistic range of signal-to-noise ratios.

@@

Another examples are DAOPHOT \citep{daophot}, and DOPHOT \citep{dophot}
which both assume analytic models for the stellar PSF profiles with centroid
coordinates being free parameters of these models.
DAOPHOT (DoPHOT) finds the centroid by fitting a Gaussian (power law) PSF to
the light profile of stars.

In this paper, we study how optimal various techniques are at centroiding 
faint stars. That is, we apply a number of centroiding methods 
to a large number of simulated faint stars with different signal-to-noise ratio and
 size realizations. Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

@@

In this work, we focus on four centroiding methods. The first method
estimates the centroid of a star by a fitting a 2d polynomial to the 3$\times$3 patch around 
the brightest pixel of the image. The second method centroids stars by
 smoothing the image of stars by a Gaussian kernel of a fixed size,
 and then applying the same 3$\times$3 polynomial trick to the smooth
 image. This method is fast and does not require any knowledge of the 
PSF. In the third method, the 3$\times$3 polynomial centroiding is applied to the
 image of star convolved with the full PSF postage stamp. The last method we consider,
 is fitting a PSF profile, assuming that we have a good PSF estimate, to the star.
%We compare the accuracy of this method with, the naive 3$\times$3 second-order 
%polynomial method, the 3$\times$3 quartic approximation
%method used in SDSS photometric pipeline, and fitting a PSF model to the 
%star.

This paper is structured as follows. In section \ref{sec:method},
we give a brief overview of the centroiding methods used in our investigation.
In section \ref{sec:data}, we discuss the simulated data, and the Cramer-Rao bound
 on centroiding error. In section \ref{sec:results}, we compare the performances of the methods
discussed in \ref{sec:method}, with the CRLB derived in \ref{sec:data}. Finally,
 we discuss and conclude in \ref{sec:discussion}.

\section{Methods}\label{sec:Centroiding methods}

\subsection{Non PSF-based methods}

\item {\bf 3$\times$3 polynomial centroiding}\quad

In order to find the centroid of a star, 
we fit a simple 2d polynomial $P(x,y)=a+bx+cy+dx^2+exy+fy^2$ to the
$3\times3$ patch centered on the brightest pixel of the
image. We apply this method to star images in the sky-dominated regime.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters of the model $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

It is important to note that the algebraic operation in (\ref{center}) involves 
inverting a 2$\times$2 curvature matrix
\beq
  D = 
  \begin{bmatrix}
      2d & e\\
      e & 2f\\
  \end{bmatrix}.
\eeq
When $D$ has a zero (or very close to zero) deteminant,
centroid estimated by (\ref{center}) gets arbitrarily 
large in which case the centroiding method leads to a catastrophic 
outlier. In order to tackle this issue, we add a soft regularization term
proportional to $\sigma^{2}$ to the determinant of $D$.

\item {\bf 3$\times$3 modified polynomial method I}\quad

Now, we modify the simple polynomial method explained above,
by smoothing the image of the star with a Gaussian kernel
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, the full-width at half-maximum of the Gaussian kernel is held at
a fixed value of 2.8 pixels (corresponding to w $\simeq$ 1.2 pixels).

Then we apply the same polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star.

\subsection{PSF-based methods}

\item {\bf 3$\times$3 modified polynomial method II}\quad

Now let us consider the case in which we have a good estimate of the PSF at
the position of the faint star under consideration. Therefore, we 
can modify the smoothing step in the following way. Instead of smoothing
with a Gaussian kernel of a fixed size, we convolve the image of star
with the full postage stamp of the PSF. Then, we apply the same polynomial
fitting to the 3$\times$3 patch around the brightest pixel of the convolved image.

The procedure of convolving the image of star with the PSF results in a
more well-sampled image, which eventually, leads to having more
information regarding the centroid of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provide a better fit 
since convolution with the PSF makes the variation of the brightness of the image
 across the 3$\times$3 patch very smooth.

\item {\bf PSF profile fitting}\quad

In addition to the methods mentioned so far, we examine fitting an exact
 PSF profile to the stars. That is, in our Cramer-Rao bound saturation tests, we find the best
 estimates of flux and centroid by optimizing the $\chi^{2}$.
 We expect this method to perform better in determing the centroid
 of stars. In theory, this method should always saturate the CRLB.

\section{Simulations and Cramer-Rao saturation tests}\label{sec:data}

In order to test the accuracy of the method mentioned in section \ref{sec:method},
 and to compare their performance at saturating the Cramer-Rao bound against each other, 
we need to simulate a large set of stars for which we know the exact position of centroids.
Furthermore, we uniformly draw the centroids of these simulated stars within sub-pixel regions. 
We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian profile. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless paramter, and $\alpha$ is
the length scale of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. 
We aim to investigate the performance of centroiding methods introduced
in section \ref{sec:method} for different background noise levels, and also different
values of $\gamma$. For further simplicity, we set the flux of all stars in our
simulations to unity. Per pixel uncertainties are assumed to be uncorrelated Gaussian,
and we only study the sky-dominated regime.

@@

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in throughout our investigation.
 In the case of our simulations, SNR given in (\ref{snr}) can be analytically
 expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We hold $\beta$ fixed at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

@@

We aim to study how centroiding errors from different methods behave with
changing SNR and size, and how close they get to saturating the CRLB. 
The closer an estimator is to saturating the CRLB, the more information
 about the quantity that we 
need to estimate is preserved. A measure for closeness of an estimator to 
saturating the CRLB is comparison of the root-mean-squared-error of that estimator 
with the CRLB.

@@

@@

\section{Results}\label{sec:results}

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], and [\ref{4}]. All methods deliver results
with RMSE larger for fainter stars.
In case of 3$\times$3 polynomial centroiding [\ref{1}], RMSE becomes
quite large as we move towards fainter stars in our simulation.
For stars with larger $\gamma$, centroid estimates from the naive 3$\times$3
 polynomial method show large deviation from the CRLB. By increasing $\gamma$,
RMSE deviates further away from the CRLB.

However, RMSE from the modified centroiding method I [\ref{2}],
is very close to the CRLB except at the very small values of SNR (SNR $<$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. When we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB. For stars with $\gamma$ = 2 pixels, 
the rate at which the RMSE from this method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. The reason for this is that even after smoothing
the with a Gaussian kernel, the image is still relatively undersampled, and not smooth enough
for a 2d polynomial fitting to provide highly accurate centroiding estimates.
For stars with $\gamma=2.8$ pixels, since the FWHM of the Gaussian kernel matches that 
of the PSF of underlying simulations, the smooth images are well-sampled and therefore, 
the method is able to deliver estimates extremely close to saturating the CRLB. As $\gamma$
gets higher for the simulations, the convolved images are not as well-sampled as those 
whose $\gamma$ matches the FWHM of the smoothing kernel and as a result, we loose
some information by fitting a polynomial to a 3$\times$3 patch. Therefore,
 we observe slight deviation from the CRLB for stars with $\gamma$ = 4 pixels, and 
slightly more deviation as we increase $\gamma$ to 5.6 pixels.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{4}]
lies on the CRLB except for simulations with SNR = 5 where the RMSE gets slightly pulled away
from the CRLB due to presence of a few outliers. Figure [\ref{3}] demonstrates that even the modified centroiding
 method II is able deliver centroiding estimates as accurate as PSF profile fitting, with the exception
of simulated stars with $\gamma$ = 2 pixels. For stars with $\gamma$ = 2 pixels, although the 3$\times$3 polynomial
fitting gets close to saturating the CRLB, it fails to saturate the CRLB since the PSF is undersampled. For simulated images with higher $\gamma$, convolving the
data with the PSF result in images that are well-sampled around the brightest pixel. This allows the 
polynomial centroiding to deliver highly accurate results that can saturate the CRLB for simulated
data with SNR $\ga$ 10.

@@

The centroid estimates obtained from the naive 3$\times$3 polynomial method [\ref{5}]
result in RMSE much larger than the CRLB in all ranges of FWHM and for all four
values of SNR in this experiment. In the case of extremely faint stars (SNR = 5),
the polynomial method is not able to deliver any reliable estimate, and completely fails.

The 3$\times$3 modified polynomial method I [\ref{6}], results in RMSE very close to the CRLB.
For all four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since the method starts to perform
slightly better as we move away from undersampled stars, and as the FWHM of the smoothing kernel
gets closer to that of the simulated images.
After approximately 3 pixels, increasing $\gamma$ result in deviation of RMSE of the method from the CRLB.
This is a characteristic of 3$\times$3 method as we apply it to a smoothed image which is still not sufficiently
well-sampled.  Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) become
closer to the CRLB.

Once again, the RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB except at SNR = 5 where the RMSE slightly deviates
from the CRLB due to presence of a few outliers [\ref{8}].
. Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by the 3$\times$3 modified polynomial method II saturate the CRLB with the
exception of simulated stars with SNR = 5, or $\gamma$ very close to 2 pixels [\ref{7}].

@@

An optimal stellar centroiding algorithm must saturate the fundamental Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroid of stars. In practice however,
this is only achievable when we have a reasonably good estimate of the PSF. Since, we do not always 
know the exact PSF profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods get close to saturating the CRLB for relatively low signal-to-noise ratio
stars.

We focused on examples from two classes of centroiding algorithms. The first class contains fast and approximate
methods that do not require any knowledge of the PSF at the position of stars. Of methods that belong to this class,
we consider a simple method for centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images, once smoothed by a Gaussian kernel of fixed width, and once without any smoothing step. 
The second class of centroiding algorithms contains methods that require knowledge of the PSF (or having a good estimate
of the PSF) at the position of stars. We considered two examples from this class. The first example
was the 3$\times$3 polynomial centroiding applied to the image of the star convolved with the PSF, and the
second example is the PSF profile fitting. In terms of saturating the Cramer-Rao bound, we compared 
the performances of these methods against each other. Throughout this work, we called the 3$\times$3 polynomial method applied 
to the image convolved with a fixed Gaussian kernel ( convolved with the PSF), 3$\times$3 modified polynomial method I (II).

%Furthermore, we compared the performance of these methods with the naive 3$\times$3 polynomial
%centroiding, 3$\times$3 quartic approximation, moment-based method, and PSF fitting.
%In order to make this comparison, we performed some tests on a large number of
%simulated data, and investigated to see if these estimators can saturate the Cramer-Rao
%lower bound, or how close to they get to saturating the bound. 
Our results suggest that in all ranges of FWHM and SNR, the PSF fitting method returns 
centroid estimates that saturate the CRLB; with the exception of SNR being less than 10, 
in which a few centroiding ouliers caused the RMSE to slightly, but not significantly, deviate
from the CRLB. We found that the estimates found by 3$\times$3 polynomial method are far away from
saturating the CRLB. On the other hand, the RMSE of centroid estimates of 
3$\times$3 modified polynomial method I are much closer to saturating the CRLB.
The performance of this method however, is limited by
two important factors: \emph{(i)} signal-to-noise ratio, and \emph{(ii)} PSF size.

The 3$\times$3 modified polynomial method I technique only takes advantage of the information
 contained in a 3$\times$3 patch around the brighest pixel of the smoothed image which is only
 well-sampled when the FWHM of the simulated image of star
 matches that of the smoothing kernel. Thus, when we apply 
this method to find the centroid of stars with larger FWHM, certain amount of
information (encoded in the Cramer-Rao lower bound) is lost, and therefore, the RMSE
of these methods deviates from the CRLB. This deviation becomes even larger at lower
signal-to-noise ratios. Besides, the performance of this method slightly degrades
in the case of undersampled stars (with FWHM close two 2 pixels). 
Presence of noise is another limitting factor.
Although this method is able to get close to saturating the CRLB in a wide range of
signal-to-noise ratios, It is not reliable in the case of extremely faint stars.
 This is partly due to the fact that in the presence of noise, 
the brightest pixel of image does not necessarily contain the centroid of stars even after smoothing 
the image.

>>>>

A common practice in astronomy is taking imaging data, and then finding the coordinates
of various light sources across the sky. Finding accuarate estimates of the centers of point
sources, convolved with telescope point spread function (and atmospheric PSF in case of
ground based telescopes), and also pixel response function, is crucial to further steps of
astronomical image processing. For instance, proper measurement of the shapes of galaxies
requires interpolating the PSF from the positions of sparsely located stars across the
image to the positions of galaxies. The accuracy of the PSF estimation therefore,
relies on how accurately we know the centroids of stars.

Ideally, we want a centroiding procedure that provides measurements as accurate as possible,
without putting a huge computational burden on the photometric pipeline.
Reducing the computational cost becomes even more important in large surveys,
where we want to estimate the centroids of thousands of point sources detected
on the telescope's focal plane.

Reaching the optimal measurement of the centroids of stars however, is limited
by the lack of knowledge about the exact shape of the PSF, and also presence of noise;
both sky noise and CCD readout noise. Thus, of particular interest is devising a fast method that returns
optimal estimates of centroids in a realistic range of signal-to-noise ratios.

@@

Another examples are DAOPHOT \citep{daophot}, and DOPHOT \citep{dophot}
which both assume analytic models for the stellar PSF profiles with centroid
coordinates being free parameters of these models.
DAOPHOT (DoPHOT) finds the centroids by fitting a Gaussian (a truncated power series for a Gaussian) PSF to
the light profile of stars.

In this paper, we study how optimal various techniques are at centroiding 
faint unsaturated stars. That is, we apply a number of centroiding methods 
to a large number of simulated faint stars with different signal-to-noise ratio and
 size realizations. Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

@@

In this work, we focus on four centroiding methods. The first method
involves fitting a PSF profile, assuming that we have a good PSF estimate, to the star. 
The second method estimates the centroid of a star by fitting a 2d polynomial to
 the 3$\times$3 patch around the brightest pixel of the image convolved with the PSF. 
The third method centroids stars by
 smoothing the image of stars by a Gaussian kernel of a fixed size,
 and then applying the same 3$\times$3 polynomial trick to the smooth
 image. This method is fast and does not require any knowledge of the 
PSF. The last method we consider, is the center of light centroiding applied to the
 7$\times$7 patch around the brightest pixel of the image.

This paper is structured as follows. In section \ref{sec:CRLB},
we discuss the Cramer-Rao lower bound, and we derive
an analytic expression for the lower bound on centroiding error
of the simulated data, and also PSF profile fitting. 
In section \ref{sec:method} we give a brief overview of approximate 
centroiding methods used in our investigation.
In section \ref{sec:data} we discuss the Cramer-Rao lower bound satuaration
tests and their corresponding simulated data.
In section \ref{sec:result}, we compare the performances of the methods
discussed in \ref{sec:method}, with the CRLB derived in \ref{sec:data}. Finally,
 we discuss and conclude in \ref{sec:discussion}.               
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Cramer-Rao lower Bound %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cramer-Rao lower bound on centroiding error}\label{sec:CRLB}

In order to test the accuracy of various centroiding methods, we need to to compare their 
performances against each other at saturating the Cramer-Rao lower bound. The closer an estimator is to saturating 
the CRLB, the more information about the quantity that we 
need to estimate is preserved. A measure for closeness of an estimator to 
saturating the CRLB is comparison of the root-mean-squared-error of that estimator 
with the CRLB.

This requires simulation of a large set of stars for which we know the exact position of centroids
and the corresponding lower bound for each star, and comparing the centroiding error with the lower bound.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THIS GOES TO TESTS%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Furthermore, we uniformly draw the centroids of these simulated stars within sub-pixel regions. 
%We aim to study how centroiding errors from different methods behave with
%changing SNR and size, and how close they get to saturating the CRLB. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian profile. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless paramter, and $\alpha$ is
the length scale of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. 
We aim to investigate the performance of centroiding methods for
 different background noise levels, and also different
values of $\gamma$. For further simplicity, we set the flux of all stars in our
simulations to unity. Per pixel uncertainties are assumed to be uncorrelated Gaussian,
and we only study the sky-dominated regime.

@@

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in throughout our investigation.
 In the case of Moffat profiles, SNR given in (\ref{snr}) can be analytically 
expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We hold $\beta$ fixed at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%SAY SOMETHING MAARRRR%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We examine fitting an exact
 PSF profile to the stars. That is, in our Cramer-Rao bound saturation tests, we find the best
 estimates of flux and centroid by optimizing the $\chi^{2}$.
 We expect this method to perform better in determing the centroids
 of stars. In theory, this method should always saturate the CRLB.

\section{Approximate methods}\label{sec:method}

\item{\bf  PSF-based 3$\times$3 polynomial centroiding}\quad Let us consider the case 
in which we have a good estimate of the PSF at
the position of the faint star under consideration. 
We can smooth the image of the star, by convolving it with the 
full postage stamp of the PSF. Then, we fit a simple 2d polynomial 
$P(x,y)=a+bx+cy+dx^2+exy+fy^2$ 
to the 3$\times$3 patch around the brightest pixel of the convolved image.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

It is important to note that the algebraic operation in (\ref{center}) involves 
inverting a 2$\times$2 curvature matrix
\beq
  D = 
  \begin{bmatrix}
      2d & e\\
      e & 2f\\
  \end{bmatrix}.
\eeq

When $D$ has a zero (or very close to zero) deteminant,
centroid estimated by (\ref{center}) gets arbitrarily 
large in which case the centroiding method leads to a catastrophic 
outlier. In order to tackle this issue, we add a soft regularization term
proportional to $\sigma^{2}$ to the determinant of $D$.

The procedure of convolving the image of star with the PSF results in a
more well-sampled image, which eventually, leads to having more
information regarding the centroids of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provide a better fit 
since convolution with the PSF makes the variation of the brightness of the image
 across the 3$\times$3 patch very smooth.

\item{\bf 3$\times$3 polynomial centroiding}\quad In the case that 
we do not have know the PSF at the position of star, we change 
the smoothing step in the following way. Instead of smoothing the image 
with the PSF, we smooth it with a Gaussian kernel of a fixed size. 
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, the full-width at half-maximum of the Gaussian kernel is held at
a fixed value of 2.8 pixels (corresponding to w $\simeq$ 1.2 pixels).

Then we apply the same 2d polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star. This is inspired by the 3$\times$3 quartic approximation 
used in the \textsl{Sloan Digital Sky Surveys} photometric pipeline \citep{sdss}.

\item {\bf Center of light centroiding}\quad In addition to the methods 
mentioned so far, we examine centroiding stars by computing their zeroth moments
in a 7$\times$7 patch around the brightest pixel of the image.
In terms of saturating the Cramer-Rao lower bound, we expect this simple 
center of light centroiding to perform worse than all other methods mentioned in 
this section. Hereafter, we call this method $7\times7$ moment centroiding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%THIS METHOD SUCKS BALLS%%%%%%%%%%%%%%%%%%%%%%%

\section{Tests}\label{sec:data}

@@

\section{Results}\label{sec:result}

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], [\ref{4}]. All methods deliver results with RMSE larger 
for fainter stars.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{1}]
lies on the CRLB except for simulations with SNR $\la$ 10 where the RMSE gets slightly pulled away
from the CRLB due to presence of a few outliers. Figure [\ref{2}] demonstrates that even the PSF-based 3$\times$3 polynomial 
centroiding is able deliver centroiding estimates as accurate as PSF profile fitting, with the exception
of simulated stars with $\gamma$ = 2 pixels. For stars with $\gamma$ = 2 pixels, although the 3$\times$3 polynomial
fitting gets close to saturating the CRLB, it fails to saturate the CRLB since the PSF is undersampled. For simulated images with higher $\gamma$, convolving the
data with the PSF result in images that are well-sampled around the brightest pixel. This allows the 
polynomial centroiding to deliver highly accurate results that can saturate the CRLB for simulated
data with SNR $\ga$ 10.

RMSE from the 3$\times$3 polynomial centroiding [\ref{3}],
is very close to the CRLB except at the very small values of SNR (SNR $\la$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. When we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB. For stars with $\gamma$ = 2 pixels, 
the rate at which the RMSE from this method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. The reason for this is that even after smoothing
the with a Gaussian kernel, the image is still relatively undersampled, and not smooth enough
for a 2d polynomial fitting to provide highly accurate centroiding estimates.
For stars with $\gamma=2.8$ pixels, since the FWHM of the Gaussian kernel matches that 
of the PSF of underlying simulations, the smooth images are well-sampled and therefore, 
the method is able to deliver estimates extremely close to saturating the CRLB. As $\gamma$
gets higher for the simulations, the convolved images are not as well-sampled as those 
whose $\gamma$ matches the FWHM of the smoothing kernel and as a result, we loose
some information by fitting a polynomial to a 3$\times$3 patch. Therefore,
 we observe slight deviation from the CRLB for stars with $\gamma$ = 4 pixels, and 
slightly more deviation as we increase $\gamma$ to 5.6 pixels.

However, in case of 7$\times$7 moment method [\ref{4}], RMSE becomes
quite large as we move towards fainter stars in our simulation.
For stars with larger $\gamma$, centroid estimates from the naive center of light 
centroiding show large deviation from saturating the CRLB. By increasing $\gamma$,
RMSE deviates further away from the CRLB.

@@

Once again, the RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB except at SNR = 5 where the RMSE slightly deviates
from the CRLB due to presence of a few outliers [\ref{5}]. 
Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by the PSF-based 3$\times$3 polynomial centroiding saturate the CRLB with the
exception of simulated stars with SNR = 5, or $\gamma$ very close to 2 pixels [\ref{6}].

The 3$\times$3 polynomial method [\ref{7}], results in RMSE very close to the CRLB.
For all four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since the method starts to perform
slightly better as we move away from undersampled stars, and as the FWHM of the smoothing kernel
gets closer to that of the simulated images.
After approximately 3 pixels, increasing $\gamma$ result in deviation of RMSE of the method from the CRLB.
This is a characteristic of 3$\times$3 method as we apply it to a smoothed image which is still not sufficiently
well-sampled.  Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) become
closer to the CRLB.

The centroid estimates obtained from the naive 7$\times$7 moment method [\ref{8}]
result in RMSE much larger than the CRLB in all ranges of FWHM and for all four
values of SNR in this experiment. In the case of extremely faint stars (SNR = 5),
the polynomial method is not able to deliver any reliable estimate, and completely fails.

@@

An optimal stellar centroiding algorithm must saturate the fundamental Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroids of stars. In practice however,
this is only achievable when we have a reasonably good estimate of the PSF. Since, we do not always 
know the exact PSF profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods come close to saturating the CRLB for relatively low signal-to-noise ratio
unsaturated stars.

We focused on examples from two classes of centroiding algorithms. The first class contains fast and approximate
methods that do not require any knowledge of the PSF at the positions of stars. Of methods that belong to this class,
we consider centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images smoothed by a Gaussian kernel of fixed width, and finding the center of light 
of a 7$\times$7 patch around the brightest pixel of the star.
The second class of centroiding algorithms contains methods that require knowledge of the PSF (or having a good estimate
of the PSF) at the positions of stars. We considered two examples from this class. The first example
is the 3$\times$3 polynomial centroiding applied to the image of the star convolved with the PSF, and the
second example is the full PSF profile fitting. In terms of saturating the Cramer-Rao bound, we compared 
the performances of these methods against each other.

%Furthermore, we compared the performance of these methods with the naive 3$\times$3 polynomial
%centroiding, 3$\times$3 quartic approximation, moment-based method, and PSF fitting.
%In order to make this comparison, we performed some tests on a large number of
%simulated data, and investigated to see if these estimators can saturate the Cramer-Rao
%lower bound, or how close to they get to saturating the bound.

Our results suggest that in all ranges of FWHM and SNR, the PSF fitting method returns 
centroid estimates that saturate the CRLB; with the exception of SNR being less than 10, 
in which case a few centroiding ouliers caused the RMSE to slightly, but not significantly, deviate
from the CRLB. We found that the estimates found by 7$\times$7 moment method, except in the case of
very high SNR values and very small ranges of $\gamma$, do not come close to
saturating the CRLB. In a considerable range of PSF sizes and background noise levels, 
this method completely fails to deliver any reliable centroiding estimates.

On the other hand, the RMSE of centroid estimates of 
3$\times$3 polynomial centroiding are much closer to saturating the CRLB in all ranges of signal-to-noise ratio.
The performance of this method however, is limited by
two important factors: \emph{(i)} signal-to-noise ratio, and \emph{(ii)} PSF size.
The 3$\times$3 polynomial centroiding technique only takes advantage of the information
 contained in a 3$\times$3 patch around the brighest pixel of the smoothed image which is only
 well-sampled when the FWHM of the simulated image of star
 matches that of the smoothing kernel. Thus, when we apply 
this method to find the centroid of stars with larger FWHM, certain amount of
information (encoded in the Cramer-Rao lower bound) is lost, and therefore, the RMSE
of these methods deviates from the CRLB. This deviation becomes even larger at lower
signal-to-noise ratios. Besides, the performance of this method slightly degrades
in the case of undersampled stars (with FWHM close two 2 pixels). 
Presence of noise is another limitting factor.
Although this method is able to get very close to saturating the CRLB in a wide range of
signal-to-noise ratios, it is not reliable in the case of extremely faint stars.
 This is partly due to the fact that in the presence of noise, 
the brightest pixel of image does not necessarily contain the centroid of stars even after smoothing 
the image.

Once we modify the polynomial method further by smoothing the image with the full postage stamp
of the PSF, we obtain results that saturate the CRLB in a wide range of PSF sizes and noise levels.  
This is due to the fact that once the images are convolved with the correct PSF, they become so well-sampled and smooth
that fitting a 2d polynomial to the 3$\times$3 patch around the brightest pixel of the smooth image is sufficient for us to obtain
results as accurate as those from fitting a PSF profile. The only conditions in which slight deviation from
saturating the CRLB occures are very low signal-to-noise ratios (SNR $\la$ 10), and undersampled PSF (FWHM $\sim$ 2 pixels).
One advantage of this method over PSF fitting method is that it is fast, and that it is able to saturate the CRLB in 
a wide range of conditions. In the case that we have a good estimate of the PSF, the 
PSF-based 3$\times$3 polynomial centroiding is a significantly much faster algorithm. 
Although this method is more accurate than 3$\times$3 polynomial centroiding, it is
slightly slower because it requires convolution of the image with the full potage stamp of the PSF.

In this investigation we showed that PSF fitting is always better at centroiding stars.
However, this method has its own disadvantages. First, we do not always know the the exact PSF. Second, finding
 the centroid by profile fitting is computationally expensive, whereas by employing any of the 3$\times$3 polynomial techniques 
considered in this study into large scale astronomical surveys can considerably reduce the computational cost
of initial astrometry of the point sources detected by these surveys.
